\chapter{Introduction}\label{ch:Introduction}
\glijbaantje{It is important to draw wisdom from many different places.\\If you take it from only one place, it becomes rigid and stale.}{Uncle Iroh}

% TODO: talk about machine learning and the advent of neuromorphic computing and the motivation for this research

\section{Reservoir computing}\label{sec:1:RC}\indexlabel{reservoir computing}
\subsection{Motivation}
\subsection{Origin}
\subsubsection{ANN}
\paragraph{Feedforward NN}
\paragraph{Recurrent NN}
%This dynamical system can be modelled as a recurrent neural network (RNN).
\subsubsection{LSM}
\subsubsection{ESN}
\subsubsection{Reservoir computing}
Reservoir computing is a machine learning framework where a ``reservoir'' --- an unspecified non-linear dynamical system --- maps input data into a higher-dimensional space, which facilitates the separation of said data by a linear transformation~\cite{KUR-24}.
A weighted sum of the components of the reservoir's state vector then provides the output of the system as a whole: this is called a \xlabel{single-layer perceptron}.
These weights can be changed in order to produce a desired output, which is referred to as ``training'' the reservoir.
Crucially, \textit{the properties of the reservoir itself are not modified during training} --- only the perceptron is. \par
Since the reservoir acts as a ``black box'', its properties can not be changed to obtain the desired response\footnote{
	It is possible to adjust a physical reservoir to exhibit the desired properties for a given task~\cite{AdaptiveProgrammableRC,gartside2022reconfigurable}, but this is different from mathematically tuning the weights of the perceptron via a simple procedure like linear regression.
}.
Therefore, it is beneficial for the reservoir to be a system with short-term memory and high dimensionality~\cite{NeuromorphicAFMspintronics,RC_RecentAdvances}.
Conveniently, these are properties exhibited by many physical systems~\cite{RC_DipoleNanomagnets,RC_PassiveFrustratedNM,RC_ASI,RC_RecentAdvances,NeuromorphicOscillators,VowelRecognition4STO,RC_DiffusiveMemristors,RC_MemristorTemporal,gartside2022reconfigurable}.
This allows the usage of physical systems without requiring higher mathematical abstractions --- apart from the single-layer perceptron at the end --- that are for example required for calculating the output of extensive neural networks.
When using a physical reservoir to perform a temporal task, the physics of the system are most effectively harnessed when the timescale of system dynamics matches that of incoming data patterns~\cite{KUR-24}.

\paragraph{Mathematical description}
% TODO: figure with right symbols
In the reservoir computing paradigm, an input signal $u(t)$ perturbs a reservoir $\mathcal{R}$, which can be any non-linear dynamical system, be it physical or abstract.
The state of the reservoir following this perturbation can then be represented by a response vector $\vc{r}(t)$.
The final output $\hat{y}(t)$ of the system is then obtained as a weighted sum $\hat{y}(t) = \vc{r}^\mathrm{T}(t) \cdot \vc{w}$ of the components of this vector $\vc{r}(t)$ --- such a weighted sum is also referred to as a single-layer perceptron.
Depending on the task performed, one or multiple perceptrons can be trained, whose interpretation then depends on the context of the specific task.
% Here, we consider the training of just a single perceptron.
When the desired input-output relation $s(t) \mapsto y(t)$ is known for a limited set of data --- the training set --- the weights $\vc{w}$ can be adjusted to minimise the difference between the output $\hat{y}(t)$ and the desired output $y(t)$.
This is usually done by ordinary least-squares (OLS) regression, which requires only a single matrix inversion.
The optimal weights then remain fixed, and the performance of the reservoir can be evaluated by feeding it the test set --- a previously unknown, yet similar data set as the training set.

% TODO: should I explain the mathematics of the matrix inversion here? Can take inspiration from~\cite{RC_NNN} eq. 4--7.
% Training of the weights $\vc{w}$ via linear regression: \hat{y}_i = x_i^T \cdot w + \epsilon_i or in matrix form y = X w + \epsilon. The OLS algorithm adjusts the weight vector w to minimize the MSE between the estimated output and the target function.

\subsubsection{Multi-reservoir techniques}
\paragraph{Single dynamical node}
\paragraph{Rotating neurons reservoir}
\subsection{Metrics} \label{sec:1:RC_metrics}
% TODO: cite relevant papers, like \cite{NeuromorphicFewShot}: in the Methods section there is a good mathematical introduction on MG, NARMA, and MC/NL, so can use this as reference that MG is mostly memory-driven.
\subsubsection{Kernel-quality, generalisation-capability, compute quality} \label{sec:1:RC_metrics_KQ}
\subsubsection{Task-agnostic metrics: nonlinearity, memory capacity, complexity (, parity check)}
% TODO: what does PC actually measure? Explain that its calculation is similar to that of memory capacity, but that the added summation and multiplication introduces some non-linearity aspect to it.
\subsubsection{Attractors etc.} % More test-like than metric-like. NARMA, MG, Lorentz
\paragraph{Mackey-Glass}
\subsubsection{Other} % Memorisation, frequency generation, classification
ICT task~\cite{farronato2022reservoir,grollier2020neuromorphic} (simple image classification) % Tasks similar to ICT are performed in those refs
\subsection{Applications or tasks}
Blood glucose level monitoring~\cite{FewMoleculeReservoir}.
\subsubsection{Digital signal processing} % Chaotic time series, speech recognition, TODO
\subsection{Physical RC}
\subsubsection{Why physical systems can be used for RC}
\subsubsection{Physical platforms suitable for RC}
\paragraph{Magnetic} % Rings and ASI
% TODO: consider adding references 308-314 of Pieter's PhD thesis, which concern RC using nanomagnets
In the context of the \spinengine project, three types of magnetic systems were considered.
These are the magnetic nanorings~\cite{DynamicEmergence_NanomagneticSystem}, focused on by the University of Sheffield, in-plane (IP) artificial spin ice (ASI)~\cite{RC_ASI}, researched by NTNU, and out-of-plane (OOP) ASI~\cite{KUR-24}, the primary interest of ETHZ.
Meanwhile, Ghent University provided simulation support for these magnetic systems.
All three present their own advantages and challenges. \par
The nanoring ensembles show promising RC performance, but did not provide straightforward on-chip input and readout methods as would be desirable for applications.
Using anisotropic magnetoresistance (AMR) only provides a single readout value --- thereby obscuring a lot of the system's dynamics --- and requires the use of lock-in amplifiers~\cite{ArchitecturesNanoringRC,Vidamour2023}.
Readout using ferromagnetic resonance (FMR) can provide a higher-dimensional readout in the form of spin-wave spectra~\cite{swindells2024fingerprinting}, but requires bulky waveguides. \par
The potential for RC in IP ASI has been demonstrated numerically~\cite{RC_ASI}, but the experimental application of input and state readout is not straightforward.
An external field is often used for input, though this is undesirable for on-chip applications; instead, it appears possible to use SOT~\cite{SOT_switching_IP}.
Efficient read-out of IP ASI remains challenging, due to the symmetry of the system coupled with the discontinuous nature of the ASI. \par % TODO END: cite roadmap and include more info
Finally, for the OOP ASI efficient input (SOT) and read-out (AMR) mechanisms had been demonstrated experimentally, but their potential for RC had not.
Therefore, this thesis will fill this gap in knowledge and assess the viability of RC with OOP ASI by making use of numerical simulations.
\paragraph{Electronic}
\paragraph{Water bucket}

\section{Artificial spin ice}\label{sec:1:ASI}\indexlabel{artificial spin ice} % TODO END: decide how exactly to use these labels, which words to label like this. Perhaps a good rule of thumb is to label the list of abbreviations, for easy reference to their first occurrence in the text.
\subsection{Nanomagnet(ism)}
\subsubsection{Physics} % See masterproef for this, IP vs. OOP also
\subsection{I/O for RC}\label{sec:1:ASI_IO}
\subsubsection{Input}
\paragraph{SOT}
% For more information on symmetry breaking by in-plane field, see~\cite{SOT_Roadmap} p.29, where also 4 other methods for symmetry breaking are presented because an external field is quite impractical. \par
\cite{SOT_FM_AFM,SOTswitchingCoPt,SOT_Roadmap,vlasov2022optimal}
\paragraph{STT} % Briefly
\subsubsection{Output}
\paragraph{AHE} % Mentioned in text, so definitely explain
\cite{AHE,AHE_Culcer}
\paragraph{AMR} % Briefly
\paragraph{FMR} % Briefly
\paragraph{SHE} % Briefly
\cite{SHE}
\subsubsection{Imaging} % Imaging can also be seen as an "output", though this should be considered a class of its own.
\paragraph{MFM} % Mentioned in text, so definitely explain
\paragraph{PEEM}

