\chapter{Publications}
\markboth{Publications}{Publications} % For headers in fncychap
\renewcommand{\bibitem}[1]{\item}

\paragraph*{Papers}
\begin{itemize}
    \bibitem{MAES-24}
    \textbf{J.~Maes}, D.~De~Gusem, I.~Lateur, J.~Leliaert, A.~Kurenkov, and B.~Van Waeyenberge.
    \newblock The design, verification, and applications of Hotspice: a Monte Carlo simulator for artificial spin ice.
    \newblock \emph{Computer Physics Communications}, page 109643, 2025.

    \bibitem{KUR-24}
    A.~Kurenkov, \textbf{J.~Maes}, A.~Pac, G.~M.~Macauley, B.~Van~Waeyenberge, A.~Hrabec, and L.~J.~Heyderman.
    \newblock Perpendicular-anisotropy artificial spin ice with spontaneous ordering: a platform for neuromorphic computing with flexible timescales.
    \newblock \emph{arXiv preprint}, arXiv:\penalty0 2408.12182, 2024.
    \newblock Accepted for publication in \emph{Communications Engineering}.

    \bibitem{MOR-24}
    L.~Moreels, I.~Lateur, D.~De~Gusem, J.~Mulkers, \textbf{J.~Maes}, M.~V.~Milosevic, J.~Leliaert, and B.~Van~Waeyenberge.
    \newblock mumax+: extensible GPU-accelerated micromagnetics and beyond.
    \newblock \emph{arXiv preprint}, arXiv:\penalty0 2411.18194, 2024.
\end{itemize}

\paragraph*{Presentations}
\begin{itemize}
    \bibitem{SPINENGINE-24}
    \textbf{J.~Maes}, A.~Kurenkov, A.~Pac, G.~M.~Mcauley, B.~Van~Waeyenberge, A.~Hrabec and L.~J.~Heyderman.
    \newblock Reservoir Computing in Thermally Active Perpendicular-Anisotropy Artificial Spin Ice.
    \newblock \emph{Workshop on Magnetic Neuromorphic Computing 2024}, Trondheim (Norway), June 17--18, 2024.
\end{itemize}

\paragraph*{Other}
\begin{itemize}
    \bibitem{FASI-23}
    \textbf{J.~Maes} and B.~Van~Waeyenberge.
    \newblock Modelling of reservoir computing in out-of-plane artificial spin ice.
    \newblock \emph{Frontiers in Artificial Spin Ice 2023}, Lenzburg (Switzerland), June 14--16, 2023 (Poster).

    \bibitem{JEMS-23}
    \textbf{J.~Maes} and B.~Van~Waeyenberge.
    \newblock Modelling of reservoir computing in out-of-plane artificial spin ice.
    \newblock \emph{13th Joint European Magnetic Symposium}, Madrid (Spain), August 27--31, 2023 (Poster).

    \bibitem{MMM-23}
    \textbf{J.~Maes}, A.~Kurenkov and B.~Van~Waeyenberge.
    \newblock Modelling of Reservoir Computing in Out-of-Plane Artificial Spin Ice.
    \newblock \emph{68th Annual Conference on Magnetism and Magnetic Materials}, Dallas, Texas (USA), October 30 -- November 3, 2023 (Poster).

    \bibitem{EAI-23}
    \textbf{J.~Maes}, A.~Kurenkov and B.~Van~Waeyenberge.
    \newblock Modelling of Reservoir Computing in Out-of-Plane Artificial Spin Ice.
    \newblock \emph{Emergent Algorithmic Intelligence Workshop 2023 `Bridging Disciplines: Novel AI-driven interdisciplinary techniques and methods'}, Mainz (Germany), December 6--8, 2023 (Poster).
\end{itemize}
